---
title: "JPRAT"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# JPRAT

<!-- badges: start -->
The R package `JPRAT` is the Joint Progression of Risk Assessment Tool, which evaluates and compares all outcomes simultaneously across multiple studies and adjusts for outcome dependencies. The JPRAT is an estimation procedure that handles multiple data hierarchy using an additive logistic mixed effect model. This algorithm handles censoring with pseudo-values and functional covariate effects with splines. The description for the detailed estimation procedure is in the following paper:

  **Garcia, T. P., Marder, K., and Wang, Y. (2017). Time-varying proportional odds model for mega-analysis of clustered event times, Biostatistics, 20(1), 129-146. \
  Garcia, T.P., Wang, Y., Shoulson, I., Paulsen, J.S. and Marder, K. (2018). Disease progression in Huntington disease: An analysis of multiple longitudinal outcomes. Journal of Huntington's disease, 7, 337-344 **
<!-- badges: end -->


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tpgarcia/JPRAT")
```
## Usage
```{r usage, results="hide",  message=FALSE}

library(JPRAT)

##get number at risk table
study.names=c("cohort", "predict", "pharos");
data.file.names=c("cohort", "predict", "pharos");
event.outcome.names=c("hdage_nobase", "mcione", "tfctwo");
nonfunctional.covariate.names=c("base_age");
functional.covariate.names="CAG";
nonfunctional.covariate.value=c(40);
functional.covariate.values.of.interest=c(46, 48, 50) ;
time.points.for.prediction=seq(46, 66, by=5);
estimated.parameters.common.for.all.studies=FALSE;

number.at.risk <- compute.number.at.risk.for.HD(study.names,
                                                data.file.names,
                                                event.outcome.names,
                                                nonfunctional.covariate.names,
                                                functional.covariate.names,
                                                nonfunctional.covariate.value,
                                                functional.covariate.values.of.interest,
                                                time.points.for.prediction,
                                                estimated.parameters.common.for.all.studies
                                                )



# Check up for the time points criteria  ##
time.points.for.conditional.prediction=c(46, 51, 56);
functional.covariate.values.of.interest.ci=c(46, 51);
functional.covariate.comparisons=c(46, 51);
time.points.of.interest=c(46, 56);
time.points.of.interest.ci=c(46, 51); 

check.criteria<-criteria.time.points.for.jprat.analysis(study.names,
                                        data.file.names,
                                        time.points.for.prediction,
                                        nonfunctional.covariate.names,
                                        functional.covariate.names,
                                        nonfunctional.covariate.value,
                                        time.points.for.conditional.prediction,
                                        functional.covariate.values.of.interest,
                                        functional.covariate.values.of.interest.ci,
                                        functional.covariate.comparisons,
                                        time.points.of.interest,
                                        time.points.of.interest.ci)
```
## Example.
We applied the combination of the Cohort, Predict and Pharos studies to JPRAT algorithm.  

```{r example, eval=F, message=FALSE}
library(JPRAT)


#Input data for JPRAT estimation procedure 
study.names=c("cohort", "predict", "pharos");
data.file.names=c("cohort", "predict", "pharos");
nonfunctional.covariate.names=c("base_age");
functional.covariate.names="CAG";
othercovariate.names=c("firstyear", "lastyear")
event.outcome.names=c("hdage_nobase", "mcione", "tfctwo");
delta.names=c("delta.hdage_nobase", "delta.mcione", "delta.tfctwo");
time.points.for.prediction=seq(46, 66, by=5);
time.points.for.conditional.prediction=c(46, 51, 56);
time.points.for.conditional.prediction.toadd=c(5);
nonfunctional.covariate.value=c(40);
functional.covariate.values.of.interest=c(46, 48, 50) ;

# parameter setup to analyze data using JPRAT estimation procedure
number.of.bootstraps=100;
use.functional.beta.intercept= TRUE ;
use.functional.event.coefficients= TRUE;
use.functional.study.coefficients=TRUE;
check.study.equality=FALSE;
what.analyzed.separately="studyevent";
estimated.parameters.common.for.all.studies=FALSE;
use.bootstrap.variance=TRUE ;
estimation.when.censoring.depends.on.z=FALSE ;
write.output=TRUE;

# JPRAT estimation

jprat.estimat.results<-jprat.wrapper(study.names=study.names, data.file.names=data.file.names,                                                    nonfunctional.covariate.names=nonfunctional.covariate.names, 
                       functional.covariate.names=functional.covariate.names,    
                       othercovariate.names=othercovariate.names, event.outcome.names=event.outcome.names, 
                       delta.names=delta.names, time.points.for.prediction=time.points.for.prediction, 
                       time.points.for.conditional.prediction=time.points.for.conditional.prediction, 
                       time.points.for.conditional.prediction.toadd=time.points.for.conditional.prediction.toadd,  
                       nonfunctional.covariate.value=nonfunctional.covariate.value,   
                       functional.covariate.values.of.interest=functional.covariate.values.of.interest, 
                       number.of.bootstraps=number.of.bootstraps, 
                       use.functional.beta.intercept=use.functional.beta.intercept, 
                       use.functional.event.coefficients=use.functional.event.coefficients, 
                       use.functional.study.coefficients=use.functional.study.coefficients,
                       check.study.equality=check.study.equality,       
                       estimated.parameters.common.for.all.studies=estimated.parameters.common.for.all.studies, 
                       what.analyzed.separately=what.analyzed.separately, 
                       estimation.when.censoring.depends.on.z=estimation.when.censoring.depends.on.z,   
                       use.bootstrap.variance=use.bootstrap.variance, write.output=write.output)


# parameters to get results 
functional.covariate.values.of.interest.ci=c(46, 51);
time.points.of.interest= c(46, 56);
time.points.of.interest.ci=seq(46, 66, by=5);
which.nonfunctional.covariate.comparisons=c(1,1);
color.names=c("firebrick1", "darkgreen", "black"); ## for color.label.key
legend.names=c("Motor Diagnosis (DCL=4)", "Cognitive Impairment", "Stage II TFC")
functional.covariate.comparisons=c(46, 51);
do.plots=TRUE;
plot.confidence.intervals=TRUE;
add.number.at.risk.legend=TRUE;
ylabel.for.plots.comparing.studies="Probability"
xlabel.for.plots.comparing.studies="Age (years)"
file.name.for.analysis="test" ## figure names
show.results.description=FALSE;

## obtain default options
default.options<-default.options.for.data.setting()

## default values
use_real_data<-default.options$use_real_data
combine.data<-default.options$combine.data

## data is reformatted to get analysis results
reformatted.data.for.analysis.results<-data.reformatted.for.analysis.results(
                                            study.names, event.outcome.names,
                                            color.names, legend.names,
                                            which.nonfunctional.covariate.comparisons)

## datasets are reformatted, so that results plots and tables can be produced;
num_study<-reformatted.data.for.analysis.results$num_study;


## load results file from JPRAT algorithm
data.truth <- read.table("out_truth_real_output_iseed_1.dat",header=FALSE)
data.theta <- read.table("out_thetaest_real_output_iseed_1.dat",header=FALSE)

if(num_study > 1){
  data.combi <- read.table("out_combi_real_output_iseed_1.dat",header=FALSE)
} else {
  data.combi <- NULL
}

data.count <- read.table("out_count_real_output_iseed_1.dat",header=FALSE)
data.count.outside <- read.table("out_outside_count_real_output_iseed_1.dat",header=FALSE)
nrisk <- read.table("out_nrisk.dat",header=TRUE)

## display all results including tables and plots
results.out <- view.all.results(
  ############################
  # arguments for JPRAT     ##
  ############################
  study.names,
  data.file.names,
  nonfunctional.covariate.names,
  functional.covariate.names,
  othercovariate.names,
  event.outcome.names,
  delta.names,
  time.points.for.prediction,
  time.points.for.conditional.prediction,
  time.points.for.conditional.prediction.toadd,
  nonfunctional.covariate.value,
  functional.covariate.values.of.interest,
  #############################
  ## Input: How to analysis  ##
  #############################
  number.of.bootstraps, #=100,
  use.functional.beta.intercept,
  use.functional.event.coefficients,
  use.functional.study.coefficients,
  check.study.equality,
  estimated.parameters.common.for.all.studies,
  what.analyzed.separately,               #="none",
  estimation.when.censoring.depends.on.z, #=FALSE,
  use.bootstrap.variance,                 #=FALSE,
  ##########################
  ## For plotting results ##
  ##########################
  functional.covariate.values.of.interest.ci,
  time.points.of.interest,
  time.points.of.interest.ci,
  label.for.alpha.values.over.time=NULL,
  which.nonfunctional.covariate.comparisons, ## return nonfunctional covariate comparison
  color.names, ## returns color.labels
  legend.names, ## returns legend.labels
  functional.covariate.comparisons,
  functional.covariate.comparisons.for.sample.size=NULL,
  #############################
  # plot options             ##
  #############################
  do.plots,
  plot.confidence.intervals,
  ###########################
  # Label figures or files ##
  ###########################
  add.number.at.risk.legend,
  ylabel.for.plots.comparing.studies,
  xlabel.for.plots.comparing.studies,
  file.name.for.analysis, ## figure names
  #######################
  ## Output from JPRAT ##
  #######################
  data.truth=data.truth,
  data.theta=data.theta,
  data.combi=data.combi,
  data.count=data.count,
  data.count.outside=data.count.outside,
  nrisk=nrisk,
  data.truth.other=NULL,
  data.theta.other=NULL,
  nrisk.other=NULL,
  data.combi.other=NULL,
  s.names.use=NULL,
  s.names.other.use=NULL,
  ########################################################
  ## for computing sample sizes: users do not need them ##
  ########################################################
  type1.error=NULL,
  type2.error=NULL,
  treatment.effect=NULL,
  dropout.rate=NULL,
  #####################################
  ## Do we show results description? ##
  #####################################
  show.results.description
	)


```
