---
title: "JPRAT"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```


<!-- badges: start -->
The R package `JPRAT` is the Joint Progression of Risk Assessment Tool, which evaluates and compares all outcomes simultaneously across multiple studies and adjusts for outcome dependencies. The JPRAT is an estimation procedure that handles multiple data hierarchy using an additive logistic mixed effect model. This algorithm handles censoring with pseudo-values and functional covariate effects with splines. The description for the detailed estimation procedure is in the following paper:

  **Garcia, T. P., Marder, K., and Wang, Y. (2017). Time-varying proportional odds model for mega-analysis of clustered event times, Biostatistics, 20(1), 129-146. \
  Garcia, T.P., Wang, Y., Shoulson, I., Paulsen, J.S. and Marder, K. (2018). Disease progression in Huntington disease: An analysis of multiple longitudinal outcomes. Journal of Huntington's disease, 7, 337-344 **
<!-- badges: end -->


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("unkyunglee/JPRAT")
```
## Usage
We applied two different data sets to run JPRAT function and get results depending on whether study and event are analyzed separately.


```{r usage, echo = F,  results="hide", message=FALSE, warning=FALSE}
library(JPRAT)


## Choose option whether study and event are analyzed separately:
## Four options are available for what.analyzed.separately="studyevent", "event", "study" or "none"

     what.analyzed.separately= "none"


## Load your data in the list format: should match order of study.names
## Choose different datasets depending on your choice for what.analyzed.separately:

    if(what.analyzed.separately=="event"||what.analyzed.separately=="studyevent"){

       input.data.list=list(cohort=data_cohort, predict=data_predict, pharos=data_pharos);

    }else if(what.analyzed.separately=="none"||what.analyzed.separately=="study"){

       input.data.list=list(study1=simu_data1_model_A, study2=simu_data2_model_A);
    }




####################################################################
## Input data for JPRAT estimation procedure: parameter settings  ##
####################################################################

    if(what.analyzed.separately=="event"||what.analyzed.separately=="studyevent"){


      study.names=c("cohort", "predict", "pharos");
      othercovariate.names=c("firstyear", "lastyear");
      event.outcome.names=c("hdage_nobase", "mcione", "dep2");
      delta.names=c("delta.hdage_nobase", "delta.mcione", "delta.dep2");
      nonfunctional.covariate.names=c("base_age");
      functional.covariate.names="CAG";
      time.points.for.prediction=seq(46, 66, by=5);
      time.points.for.conditional.prediction=c(46,51, 56);
      time.points.for.conditional.prediction.toadd=c(5);
      nonfunctional.covariate.value=c(40);
      functional.covariate.values.of.interest=c(46, 48, 50) ;

    }else if(what.analyzed.separately== "none"||what.analyzed.separately=="study"){

      study.names=c("study1", "study2");
      othercovariate.names=NULL;
      event.outcome.names=c("event1", "event2");
      delta.names=c("delta.event1", "delta.event2");
      nonfunctional.covariate.names=c("base_age");
      functional.covariate.names="CAG";
      time.points.for.prediction=  c(seq(40,60,by=1));
      time.points.for.conditional.prediction=c(40, 55);
      time.points.for.conditional.prediction.toadd=c(5,10);
      nonfunctional.covariate.value=c(42);
      functional.covariate.values.of.interest=c(42);

    }


    if(what.analyzed.separately!="none"){
      check.study.equality=FALSE
      estimated.parameters.common.for.all.studies=FALSE
    }else{
      check.study.equality=TRUE
      estimated.parameters.common.for.all.studies=TRUE
    }

    number.of.bootstraps=10;
    use.functional.beta.intercept= TRUE ;
    use.functional.event.coefficients= TRUE;
    use.functional.study.coefficients=TRUE;
    use.bootstrap.variance=TRUE;
    estimation.when.censoring.depends.on.z=FALSE ;
    write.output=TRUE;

#############################################
## a function to estimate JPRAT algorithm  ##
#############################################

  jprat.estimat.results<-jprat.wrapper(study.names=study.names,
                                       input.data.list=input.data.list,
   nonfunctional.covariate.names=nonfunctional.covariate.names,
   functional.covariate.names=functional.covariate.names,
   othercovariate.names=othercovariate.names,
   event.outcome.names=event.outcome.names,
   delta.names=delta.names,
   time.points.for.prediction=time.points.for.prediction,
   time.points.for.conditional.prediction=time.points.for.conditional.prediction,
   time.points.for.conditional.prediction.toadd=time.points.for.conditional.prediction.toadd,
   nonfunctional.covariate.value=nonfunctional.covariate.value,
   functional.covariate.values.of.interest=functional.covariate.values.of.interest,
   number.of.bootstraps=number.of.bootstraps,
   use.functional.beta.intercept=use.functional.beta.intercept,
   use.functional.event.coefficients=use.functional.event.coefficients,
   use.functional.study.coefficients=use.functional.study.coefficients,
   check.study.equality=check.study.equality,
   estimated.parameters.common.for.all.studies=estimated.parameters.common.for.all.studies,
   what.analyzed.separately=what.analyzed.separately,
   estimation.when.censoring.depends.on.z=estimation.when.censoring.depends.on.z,
   use.bootstrap.variance=use.bootstrap.variance, write.output=write.output)

##########################################################
## Get a table for the number of people who are at risk ##
##########################################################

    if(what.analyzed.separately=="event"||what.analyzed.separately=="studyevent"){

      study.names=c("cohort", "predict", "pharos");
      event.outcome.names=c("hdage_nobase", "mcione", "dep2");
      nonfunctional.covariate.names=c("base_age");
      functional.covariate.names="CAG";

    }else if(what.analyzed.separately=="none"||what.analyzed.separately=="study"){

      study.names=c("study1", "study2");
      othercovariate.names=NULL;
      event.outcome.names=c("event1", "event2");
      delta.names=c("delta.event1", "delta.event2");
      nonfunctional.covariate.names=c("base_age");
      functional.covariate.names="CAG";

    }



    if(what.analyzed.separately=="event"||what.analyzed.separately=="studyevent"){

      time.points.for.prediction=seq(46, 66, by=5);
      functional.covariate.values.of.interest=c(46, 48, 50) ;
      nonfunctional.covariate.value=c(40);

    }else if(what.analyzed.separately=="none"||what.analyzed.separately=="study"){

      time.points.for.prediction=  c(seq(40,60,by=1))
      functional.covariate.values.of.interest=c(42);
      nonfunctional.covariate.value=c(42);

    }

    if(what.analyzed.separately!="none"){
      estimated.parameters.common.for.all.studies=FALSE
    }else{
      estimated.parameters.common.for.all.studies=TRUE
    }

###############################################################################
## a function to create a table for the number of patients who are at risk   ##
###############################################################################

    number.at.risk <- compute.number.at.risk.for.HD(study.names,
     input.data.list,
     event.outcome.names,
     nonfunctional.covariate.names,
     functional.covariate.names,
     nonfunctional.covariate.value,
     functional.covariate.values.of.interest,
     time.points.for.prediction,
     estimated.parameters.common.for.all.studies,
     write.output=TRUE
     )




##############################
## parameters to get results #
##############################

    if(what.analyzed.separately=="event"||what.analyzed.separately=="studyevent"){

      functional.covariate.values.of.interest.ci=c(46, 51);
      time.points.of.interest= c(46, 56);
      time.points.of.interest.ci=seq(46, 66, by=5);
      functional.covariate.comparisons=c(46, 51);
      time.points.for.prediction=seq(46, 66, by=5);
      functional.covariate.values.of.interest=c(46, 48, 50) ;
      nonfunctional.covariate.value=c(40);
      color.names=c("firebrick1", "darkgreen", "black");
      legend.names=c("Motor Diagnosis (DCL=4)");


    }else{

      functional.covariate.values.of.interest.ci=c(40, 55);
      time.points.of.interest= c(40, 55);
      time.points.of.interest.ci=seq(40, 60, by=5);# c(seq(40,60,by=1))
      functional.covariate.comparisons=c(40, 55);
      time.points.for.prediction=  c(seq(40,60,by=1))
      functional.covariate.values.of.interest=c(42);
      nonfunctional.covariate.value=c(42);
      color.names=c("firebrick1", "darkgreen");
      legend.names=c("event1", "event2");

    }


    do.plots=TRUE;
    plot.confidence.intervals=TRUE;
    add.number.at.risk.legend=TRUE;
    ylabel.for.plots.comparing.studies="Probability"
    xlabel.for.plots.comparing.studies="Age (years)"
    file.name.for.analysis="test" ## figure names
    show.results.description=TRUE;


## is.nrisk.data.frame=TRUE when users choose write.output=TRUE 
## and output for nrisk returns as the form of dataframe.
## is.nrisk.data.frame=FALSE when users choose write.output=FALSE 
## and output for nrisk returns as an array form.
## users need to choose write.jprat.output=TRUE in jprat.wrapper function; 
## FALSE otherwise.


    if(write.output==TRUE){

    is.nrisk.data.frame=TRUE;      
    write.jprat.output=TRUE;     
    
    }else{
      is.nrisk.data.frame=FALSE;
      write.jprat.output=FALSE;
    }

    
####################################
## functions to get default values #
####################################

    if(what.analyzed.separately=="studyevent"){


      study.names=c("cohort", "predict", "pharos");
      event.outcome.names=c("hdage_nobase", "mcione", "dep2");
      legend.names=c("Motor Diagnosis (DCL=4)")


    }else if(what.analyzed.separately=="event"){

      study.names=c("cohort", "predict", "pharos");
      event.outcome.names=c("hdage_nobase", "mcione", "dep2");
      legend.names=c("Motor Diagnosis (DCL=4)", "Cognitive Impairment", "Stage II TFC")

    }else if(what.analyzed.separately=="none"||what.analyzed.separately=="study"){

      color.names=c("firebrick1", "darkgreen"); ## for color.label.key
      event.outcome.names=c("event1", "event2")
      legend.names=c("Motor Diagnosis (DCL=4)", "Cognitive Impairment");

    }


    ## allocate jprat outputs
    jprat.output<-jprat.estimat.results$jprat.output
    nrisk<-number.at.risk


##################################################################
## a function to display all results including tables and plots ##
##################################################################

    results.out <- view.all.results(
      ############################
      # arguments for JPRAT     ##
      ############################
      study.names,
      #data.file.names,
      input.data.list,
      nonfunctional.covariate.names,
      functional.covariate.names,
      othercovariate.names,
      event.outcome.names,
      delta.names,
      time.points.for.prediction,
      time.points.for.conditional.prediction,
      time.points.for.conditional.prediction.toadd,
      nonfunctional.covariate.value,
      functional.covariate.values.of.interest,
      #############################
      ## Input: How to analysis  ##
      #############################
      number.of.bootstraps, #=100,
      use.functional.beta.intercept,
      use.functional.event.coefficients,
      use.functional.study.coefficients,
      check.study.equality,
      estimated.parameters.common.for.all.studies,
      what.analyzed.separately,
      estimation.when.censoring.depends.on.z,
      use.bootstrap.variance,
      ##########################
      ## For plotting results ##
      ##########################
      functional.covariate.values.of.interest.ci,
      time.points.of.interest,
      time.points.of.interest.ci,
      label.for.alpha.values.over.time=NULL,
      color.names=color.names, ## returns color.labels
      legend.names=legend.names, ## returns legend.labels
      functional.covariate.comparisons=functional.covariate.comparisons,
      functional.covariate.comparisons.for.sample.size=NULL,
      #############################
      # plot options             ##
      #############################
      do.plots=do.plots,
      plot.confidence.intervals=plot.confidence.intervals,
      ###########################
      # Label figures or files ##
      ###########################
      add.number.at.risk.legend=add.number.at.risk.legend,
      ylabel.for.plots.comparing.studies=ylabel.for.plots.comparing.studies,
      xlabel.for.plots.comparing.studies=xlabel.for.plots.comparing.studies,
      file.name.for.analysis=file.name.for.analysis, ## figure names
      #######################
      ## Output from JPRAT ##
      #######################
      jprat.output=jprat.output,
      nrisk=nrisk,
      #####################################
      ## Do we show results description? ##
      #####################################
      show.results.description=show.results.description,
      is.nrisk.data.frame=is.nrisk.data.frame,
      write.jprat.output=write.jprat.output
    )




```
